## Tumor Mutational Burden, TP53 Pathway, and Immunotherapy
```{r}
#| echo: false
#| message: false
#| warning: false

library(tidyverse)

# Read the genomics and demographics dataset
genomics_data <- read_csv("demographics_genomics_912_dataset.csv")

# Plot 1: TMB by Immunotherapy status, faceted by TP53 Pathway
genomics_data |> 
  ggplot(aes(x = `Post Sample Immunotherapy`, y = `TMB (nonsynonymous)`, fill = `Post Sample Immunotherapy`)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(width = 0.2, alpha = 0.3, color = "#2C3E50") +
  facet_wrap(~ `TP53 Pathway`, labeller = label_both) +
  scale_fill_manual(values = c("No" = "#E64B35FF", "Yes" = "#00A087FF")) +
  scale_y_log10() +
  labs(
    subtitle = "Comparing patients with and without TP53 pathway mutations",
    x = "Received Post-Sample Immunotherapy",
    y = "TMB (log10 scale)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

According to the data, we can see that of you have a higher TMB score, you are more likely to need immunotherapy. This makes sense because since TMB meausres the changes of DNA on a tumor, it should be more likely to for a patient to need a form f immunotherapy if the DNA on their tumor is very different from the DNA in their cells. From this data, we can also see that if you have a higher TMB score, you are more likely to have a TP53 pathway, meaning your body naturally fights against a tumor. This also makes sense because the more variation in the DNA of a tumor, the more likely it should be for your body to recognize that that tumor is bad and try to fight it. 

## Race, Sex, and Immunotherapy
```{r}
#| echo: false
#| message: false
#| warning: false

# Plot 2: Counts of Race Category by Sex and Immunotherapy status
genomics_data |> 
  ggplot(aes(x = `Race Category`, fill = `Post Sample Immunotherapy`)) +
  geom_bar(position = position_dodge(width = 0.8), alpha = 0.8) +
  facet_wrap(~ Sex) +
  scale_fill_manual(values = c("No" = "#3C5488FF", "Yes" = "#F39B7FFF")) +
  labs(
    subtitle = "Patient counts by race and sex, separated by post-sample immunotherapy",
    x = "Race Category",
    y = "Patient Count",
    fill = "Immunotherapy"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

This plot suggests that white females are significantly more likely to not recieve immunotherapy after surgery/biopsy than white males. We can see this through the large difference between the two bars for white females compared to white males. 

## Age and Immunotherapy
```{r}
#| echo: false
#| message: false
#| warning: false

plot_data <- genomics_data |>
  mutate(
    Immunotherapy = case_when(
      `Post Sample Immunotherapy` == "Yes" ~ "Received",
      `Post Sample Immunotherapy` == "No" ~ "Did Not Receive",
      TRUE ~ NA_character_
    )
  )

age_density <- density(plot_data$`Age at Surgery/Biopsy`)

plot_data <- plot_data |>
  mutate(
    density_score = approx(age_density$x, age_density$y, xout = `Age at Surgery/Biopsy`)$y
  )

ggplot(plot_data, aes(x = `Age at Surgery/Biopsy`, y = 0, color = Immunotherapy, alpha = density_score)) +
  geom_jitter(width = 0, height = 0.05, size = 1.8) +
  scale_color_manual(values = c("Received" = "green", "Did Not Receive" = "red")) +
  scale_alpha_continuous(range = c(0.2, 0.95)) +
  labs(
    x = "Age at Surgery/Biopsy",
    y = ""
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```
We can see from the graph, that there isn't a clear correlation between age and immunotherapy after surgery/biopsy. However, we can note that most people that did not recieve immunotherapy after surgery/biopsy were between ages 60 and 80. However, the data for people who did recieve immunotherapy after surgery/biopsy is more spread out and doesn't have a clear age range. 
