---
title: "Model"
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidymodels)
library(tidyverse)

# Load dataset
genomics_data <- read_csv("demographics_genomics_912_dataset.csv")
```

## Overview

This page presents the statistical model predicting immunotherapy use in patients with lung adenocarcinoma. Since the question is Predictive, we will use a logistic regression to model the data. 

## Table 

```{r}
#| echo: false
#| message: false
#| warning: false

logistic_reg() |>
  set_engine("glm") |> 
  fit(as.factor(`Post Sample Immunotherapy`) ~ `Age at Surgery/Biopsy` + Sex + `Race Category` + `TP53 Pathway` + `TMB (nonsynonymous)`, data = genomics_data) |>
  tidy(conf.int = TRUE)
```
According to this table, we can interpret that if you have a TP53 pathway, you are 45.2% more likely to recieve immmunotherapy after a surgery/biopsy. This estimate is statistically significant which is proven by the low p value of 0.0106 and the confidence interval (0.110 to 0.804) not including the number 0. 

## Data Generating Mechanism 

The probability $p_i$ of a patient receiving post-sample immunotherapy is modeled using logistic regression:

$$\text{logit}(p_i) = \ln\left(\frac{p_i}{1 - p_i}\right) = \beta_0 + \beta_1(\text{Age}_i) + \beta_2(\text{Sex}_{\text{Male}, i}) + \beta_3(\text{Race}_{\text{Black}, i}) + \beta_4(\text{Race}_{\text{Other}, i}) + \beta_5(\text{Race}_{\text{Unknown}, i}) + \beta_6(\text{Race}_{\text{White}, i}) + \beta_7(\text{TP53}_{\text{Yes}, i}) + \beta_8(\text{TMB}_i)$$

Using the estimated parameters from our model:

$$\text{logit}(p_i) = -0.794 - 0.0169(\text{Age}_i) - 0.0523(\text{Sex}_{\text{Male}, i}) + 0.460(\text{Race}_{\text{Black}, i}) + 0.476(\text{Race}_{\text{Other}, i}) + 0.667(\text{Race}_{\text{Unknown}, i}) + 0.251(\text{Race}_{\text{White}, i}) + 0.452(\text{TP53}_{\text{Yes}, i}) + 0.0222(\text{TMB}_i)$$

