---
title: "Model"
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(tidymodels)
library(tidyverse)

# Load dataset
genomics_data <- read_csv("demographics_genomics_912_dataset.csv")
```

## Overview

This page presents the statistical model predicting immunotherapy use in patients with lung adenocarcinoma. Since the question is Predictive, we will use a logistic regression to model the data. 

## Table 

```{r}
#| echo: false
#| message: false
#| warning: false

logistic_reg() |>
  set_engine("glm") |> 
  fit(as.factor(`Post Sample Immunotherapy`) ~ `Age at Surgery/Biopsy` + Sex + `Race Category` + `TP53 Pathway` + `TMB (nonsynonymous)`, data = genomics_data) |>
  tidy(conf.int = TRUE)
```
According to this table, we can interpret that if you have a TP53 pathway, you are 45.2% more likely to recieve immmunotherapy after a surgery/biopsy. This is proven by the low p value of 0.0106 and the confidence interval (0.110 to 0.804) not including the number 0. 

## Data Generating Mechanism 

The probability $p_i$ of a patient receiving post-sample immunotherapy is modeled using logistic regression:

$$\text{logit}(p_i) = \ln\left(\frac{p_i}{1 - p_i}\right) = \beta_0 + \beta_1(\text{Age}_i) + \beta_2(\text{Sex}_{\text{Male}, i}) + \beta_3(\text{Race}_{\text{Black}, i}) + \beta_4(\text{Race}_{\text{Other}, i}) + \beta_5(\text{Race}_{\text{Unknown}, i}) + \beta_6(\text{Race}_{\text{White}, i}) + \beta_7(\text{TP53}_{\text{Yes}, i}) + \beta_8(\text{TMB}_i)$$

Using the estimated parameters from our model:

$$\text{logit}(p_i) = -0.794 - 0.0169(\text{Age}_i) - 0.0523(\text{Sex}_{\text{Male}, i}) + 0.460(\text{Race}_{\text{Black}, i}) + 0.476(\text{Race}_{\text{Other}, i}) + 0.667(\text{Race}_{\text{Unknown}, i}) + 0.251(\text{Race}_{\text{White}, i}) + 0.452(\text{TP53}_{\text{Yes}, i}) + 0.0222(\text{TMB}_i)$$

## More Plots

```{r}
#| echo: false
#| message: false
#| warning: false

plot_data <- genomics_data |>
  mutate(
    Immunotherapy = case_when(
      `Post Sample Immunotherapy` == "Yes" ~ "Received",
      `Post Sample Immunotherapy` == "No" ~ "Did Not Receive",
      TRUE ~ NA_character_
    )
  )

age_density <- density(plot_data$`Age at Surgery/Biopsy`)

plot_data <- plot_data |>
  mutate(
    density_score = approx(age_density$x, age_density$y, xout = `Age at Surgery/Biopsy`)$y
  )

ggplot(plot_data, aes(x = `Age at Surgery/Biopsy`, y = 0, color = Immunotherapy, alpha = density_score)) +
  geom_jitter(width = 0, height = 0.05, size = 1.8) +
  scale_color_manual(values = c("Received" = "green", "Did Not Receive" = "red")) +
  scale_alpha_continuous(range = c(0.2, 0.95)) +
  labs(
    title = "Age vs. Post-Sample Immunotherapy",
    x = "Age at Surgery/Biopsy",
    y = ""
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```
We can see from the graph, that there isn't a clear correlation between age and immunotherapy after surgery/biopsy. However, we can note that most people that did not recieve immunotherapy after surgery/biopsy were between ages 60 and 80. However, the data for people who did recieve immunotherapy after surgery/biopsy is more spread out and doesn't have a clear age range. 
